---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load necessary packages
```{r}
library(ggplot2)
library(mgcv)
library(tidyverse)
library(readxl)
library(lubridate)
```

Read Fyrisån concentration and discharge data (last year). Create year, month and day variables
```{r}
fyris1<-read_excel( "C:\\Users\\claudia\\Documents\\Forskning\\Time-varying coefficients\\R code\\Model illustration\\Fyris Flottsund 2022.xlsx")%>%
  mutate(Year=Provtagningsår, Month=Provtagningsmånad, Day=Provtagningsdag)
fyrisQ2020<-read_excel( "C:\\Users\\claudia\\Documents\\Forskning\\Time-varying coefficients\\R code\\Model illustration\\Fyris Q 2020.xlsx")%>%
  mutate(Year=year(Datum), Month=month(Datum), Day=day(Datum), Q=`Total stationskorrigerad vattenföring`)%>%dplyr::select(Year, Month, Day, Q)
#fyrisQ2021<-read_excel( "C:\\Users\\claudia\\Documents\\Forskning\\Time-varying coefficients\\R code\\Model illustration\\Fyris Q 2022.xlsx")%>%
#  mutate(Year=year(Datum), Month=month(Datum), Day=day(Datum), Q=`Total stationskorrigerad vattenföring`)%>%dplyr::select(Year, Month, Day, Q)%>%filter(Year==2021)
```

Read older discharge data i.e before 1981 from the second file. Merge datasets and create a date variable
```{r}
fyrisQ70<-read_excel( "C:\\Users\\claudia\\Documents\\Forskning\\Time-varying coefficients\\R code\\FyrisQ.xlsx")%>%
  mutate(Year=ÅR, Month=MÅNAD, Day=DAG, Q=`Q_m3/s`)%>%
  filter(ÅR<1981)%>%dplyr::select(-DAG, -MÅNAD, -ÅR)

fyrisQ70%>%full_join(fyrisQ2020)%>%#full_join(fyrisQ2021)%>%
  right_join(fyris1)%>%
  mutate(Date=as.Date( paste(Month, Day , Year , sep = "." )  , format = "%m.%d.%Y"))%>%
  arrange(Date)%>%filter(Year<2021)->Fyris2
```

Create decimal date variable and date within year variable. Log-transform concentration and discharge data. Remove observations that have no discharge measurements.
```{r}
Fyris2%>%
  mutate(date=ymd(Date),
 dec.date=decimal_date(date),
 decimal_during = dec.date - year(date),
 log_q=log10(Q),
 log_PO4=log10(`PO4-P (µg/l P)`),
 log_TP=log10(`Tot-P (µg/l P)`),
 log_NO2NO3=log10(`NO2+NO3-N (µg/l N)`))%>%
 drop_na(Q)->Fyris
```

Table 3
modFyris_constant_without: model with a constant c-q slope and no concentration trend (LM_Constant)
modFyris_constant_withtrend: model with a constant c-q slope and a concentration trend included (GAM_constant)

```{r}
modFyris_constant_without<-lm(log_TP~log_q, data=Fyris)
summary(modFyris_constant_without)


Fyrismod_constant_withtrend<-gamm(log_TP~s(dec.date)+log_q, method="REML", data=Fyris)

summary(Fyrismod_constant_withtrend$gam)

```

######################################################################################################
Figure 4, left: A plot indicating the concentration- discharge relationship for different time periods
######################################################################################################

Create a factor to indicate different time spans

```{r}
Fyris%>%
 mutate(decenn2=case_when(Year<1975~"1965-1974",
                          Year >=1975 & Year<1985~"1975-1984",
                          Year>=1985 & Year <1995~"1985-1994",
                          Year>=1995 & Year <2005 ~"1995-2004",
                          Year>=2005 ~"2005-2020"))->Fyris
```

Define a color palette and plot
```{r}
library(RColorBrewer)
my_palette <- rev(brewer.pal(name="Blues",n=9)[5:9])

Figure4left<-Fyris%>%
  ggplot(aes(x=log_q, y=log_TP, col=decenn2))+
  geom_point()+
  geom_smooth(method="lm", se=FALSE,data=Fyris%>%filter(decenn2=="1965-1974"))+
  geom_smooth(method="lm", se=FALSE, data=Fyris%>%filter(decenn2=="1975-1984"))+
  geom_smooth(method="lm", se=FALSE, data=Fyris%>%filter(decenn2=="1985-1994"))+
  geom_smooth(method="lm", se=FALSE, data=Fyris%>%filter(decenn2=="1995-2004"))+
  geom_smooth(method="lm", se=FALSE, data=Fyris%>%filter(decenn2=="2005-2020"))+
  xlab("Discharge (log-transformed)")+
  ylab("Total phosphorus (log-transformed)")+
  labs(col = "Time period")+
  xlim(0,2.1)+
  theme_bw()+
  theme(text= element_text(size = 16), legend.position="bottom", axis.title.x = element_text(size = 16), axis.title.y = element_text(size = 16))+
  scale_color_manual(values=my_palette)+
  guides(color=guide_legend(nrow=2, byrow=TRUE))


```

```{r}
ggsave(filename = "Figure 4 left.jpg",    
       plot = Figure4left,
       #plot = last_plot(),
       width = 6, height = 6, device = "jpg", dpi = 300 )
```

###################################################
Figure 4, right: Time-varying c-q slopes in Fyrisån
###################################################

modfyris_GAM_T: a  model with c-q slopes that vary over years (GAM_timevarying)
Extract c-q slopes (coeff) and plot

```{r}
Fyris$logq_cent=Fyris$log_q-mean(Fyris$log_q)
modFyris_GAM_T<-gamm(log_TP~s(dec.date,bs="ts", k=40) +s(dec.date,bs="ts", by=logq_cent),data=Fyris, correlation=corCAR1(form=~dec.date), method="REML")

terms1d<-predict.gam(modFyris_GAM_T$gam, type="terms", se.fit=TRUE)

Fyris$trend<-terms1d$fit[,1]+attr(terms1d, "constant")#+terms1d$fit[,2]/Fyris$logq_cent
Fyris$trend.upper<-Fyris$trend+1.96*terms1d$se.fit[,1]
Fyris$trend.lower<-Fyris$trend-1.96*terms1d$se.fit[,1]
Fyris$coeff<-terms1d$fit[,2]/Fyris$logq_cent
Fyris$coeff.upper<-Fyris$coeff+1.96*terms1d$se.fit[,2]/abs(Fyris$logq_cent)
Fyris$coeff.lower<-Fyris$coeff-1.96*terms1d$se.fit[,2]/abs(Fyris$logq_cent)

```


```{r}
gam.check(modFyris_GAM_T$gam)
concurvity(modFyris_GAM_T$gam)
```



Plot c-q slopes.
```{r}
Figure4right<-ggplot(Fyris, aes(dec.date, coeff,coeff.lower,coeff.upper)) +
  geom_line(aes(y = coeff), color="red", lwd=2)+
  geom_line(aes(y = coeff.lower))+
  geom_line(aes(y = coeff.upper))+
  ylab("c-q slope")+
  xlab("Date")+
  geom_ribbon(aes(ymin=coeff.lower, ymax=coeff.upper),alpha=0.09, fill = "blue")+
  theme_bw()+
  theme(text= element_text(size = 16))
```
```{r}
ggsave(filename = "Figure 4 right.jpg",    
       plot = Figure4right,
       #plot = last_plot(),
       width = 6, height = 6, device = "jpg", dpi = 300 )
```


Modelled trend (not shown)
```{r}

ggplot(Fyris, aes(dec.date, trend)) +
  geom_line(aes(y = trend), color="red", lwd=2)+
  geom_line(aes(y = trend.lower))+
  geom_line(aes(y = trend.upper))+
  ylab("Modeled trend")+
  xlab("Date")+
  geom_ribbon(aes(ymin=trend.lower, ymax=trend.upper),alpha=0.09, fill = "blue")+
  theme_bw()+
  theme(text= element_text(size = 16))
```





Figure S1: Trend in TP in the Fyriså catchment

```{r}
fyris_trend_only<-gamm(log_TP~s(dec.date, k=60),correlation=corCAR1(form = ~dec.date),method="REML", data=Fyris)
terms<-predict.gam(fyris_trend_only$gam, type="terms", newdata=Fyris, se.fit=TRUE)

Fyris$fit_only_trend<-terms$fit[,1]+attr(terms, "constant")

FigureS1<-ggplot(Fyris, aes(dec.date, log_TP,fit_only_trend)) +
  geom_point(aes(x=dec.date, y=log_TP))+
  geom_line(aes(y = fit_only_trend), color="red", lwd=2)+
  ylab("Total phosphorus (log-transformed)")+
  xlab("Date")+
  theme_bw()+
  theme(text= element_text(size = 16))
```


```{r}
ggsave(filename = "Figure S1.jpg",    
       plot = FigureS1,
       #plot = last_plot(),
       width = 6, height = 6, device = "jpg", dpi = 300 )
```

#################################################################
Figure S4: Prediction plot
##################################################################
```{r}
modFyris_GAM_T_constant_cq<-gamm(log_TP~s(dec.date,bs="ts", k=40) +logq_cent,data=Fyris, correlation=corCAR1(form=~dec.date), method="REML")

modFyris_GAM_T<-gamm(log_TP~s(dec.date,bs="ts", k=40) +s(dec.date,bs="ts", by=logq_cent),data=Fyris, correlation=corCAR1(form=~dec.date), method="REML")

Fyris$resid_const<-residuals(modFyris_GAM_T_constant_cq$gam)
Fyris$resid_T<-residuals(modFyris_GAM_T$gam)
Fyris$pred_const<-as.numeric(predict(modFyris_GAM_T_constant_cq$gam , type="response"))
Fyris$pred_T<-as.numeric(predict(modFyris_GAM_T$gam, type="response"))


Fyris%>%pivot_longer(cols=c(pred_T, pred_const))->Fyris_b

labels<- as_labeller(c(pred_const="GAM[constant]",pred_T="GAM[timevarying]"),default=label_parsed)


FigureS4middle<-Fyris_b%>%
  ggplot(aes(x=log_TP,y=value))+
  geom_point(col="blue")+
  geom_abline(slope=1, intercept=0, lwd=2)+
  facet_wrap(~name, labeller=labels)+
  ylab("Predicted values")+
  xlab("Total phosphorus (log-transformed)")+
  theme_classic()+
   theme(strip.text.x = element_text(size = 16),text= element_text(size = 16))



```

```{r}
ggsave(filename = "Figure S4 middle.jpg",    
       plot = FigureS4middle,
       #plot = last_plot(),
       width = 7, height = 5, device = "jpg", dpi = 300 )
```
