---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load necessary packages
```{r}
library(tidyverse)
library(mgcv)
library(lubridate)
library(readxl)
```

A function to change decimal comma to decimal point
```{r}
replace_comma <- function(values){
    values_parsed <- values %>% as.character() %>% parse_number(locale = locale(decimal_mark = ","))
}
```

Read concentration data för H29, replace decimal comma with decimal point, log-transform total phosphorus, create variables year, month and day
```{r}
mtransp<-read_excel("Jordbruk_analysvärden.xls", sheet="Analysvärden", skip=2, guess_max=20000)%>%filter(ID=="H29", Provtyp=="Manuellt prov")%>%
  mutate_at(c( "Totalkväve_mg/l", "Totalfosfor_mg/l"),replace_comma)%>%
  mutate(logTN=log10(`Totalkväve_mg/l`), logTP=log10(`Totalfosfor_mg/l`),Year=as.numeric(substr(Datum, 1,4)), Month=as.numeric(substr(Datum, 6,7)), Day=as.numeric(substr(Datum, 9,10)))
```

Read discharge data, replace decimal comma with decimal point, create varialbe year, month and day, remove observations with a discharge of 0. 
```{r}
disch<-read_excel("Vattenföring_H-län_210323.xls", sheet="Vattenföring", skip=2, guess_max=100000)%>%filter(ID=="H29")%>%
   mutate_at(c("Dygnsmedelflöde_l/s"),replace_comma)%>%
  mutate(logq=log10(`Dygnsmedelflöde_l/s`), Year=as.numeric(substr(Datum, 1, 4)), Month=as.numeric(substr(Datum, 6,7)), Day=as.numeric(substr(Datum, 9,10)))%>%
  filter(!(`Dygnsmedelflöde_l/s`==0))
```

Combine concentration and discharge data, create a decimal date variable and a variable indicating date within the year.
```{r}
H29<-mtransp%>%left_join(disch)%>%mutate(Date=as.Date( paste(Month, Day , Year , sep = "." )  , format = "%m.%d.%Y"), dec.date=decimal_date(Date), dec.during=dec.date-year(Date))

```
Table 3:
modH29_constant_withouttrend: model with constant c-q slope and without concentration trend (LM_Constant)
modH29_constant_withtrend: model with constant c-q slope including a concentration trend (GAM_constant)

```{r}
modH29_constant_withouttrend<-lm(logTN~logq, data=H29)
summary(modH29_constant_withouttrend)

modH29_constant_withtrend<-gamm(logTN~s(dec.date)+logq, correlation=corCAR1(form = ~dec.date),method="REML", data=H29)
summary(modH29_constant_withtrend$gam)
```


Time-varying c-q slope model, changing over seasons (GAM_time-varying and seasonal)

```{r}
modH29_GAM_TS<-gamm(logTN~te(Year, dec.during, bs=c("cr", "cc")) +te(Year, dec.during, by=logq,  bs=c("cr", "cc")),correlation=corCAR1(form=~dec.date),data=H29, method="REML")


#Without autocorrelation term,in case the upper one does not run
#modH29_GAM_TS<-gamm(logTP~te(Year, dec.during, bs=c("cr", "cc")) +te(Year, dec.during, by=logq,  bs=c("cr", "cc")),data=H29, method="REML")
```

Extract fitted values and c-q slope for the entire year
```{r}
terms29<-predict.gam(modH29_GAM_TS$gam, type="terms", newdata=H29, se.fit=TRUE)

```

#########################################################
Figure 5, left: Plot of the year-season varying c-q slope
#########################################################

Define a grid for plotting, years from start to end, season as decimal 0 to 1, logq equal to 1 to extract the slope parameter. Estimated coefficients on this grid and plot
```{r}
new<-expand.grid(dec.during=seq(0,1, length.out=50), logq=rep(1, 50), Year=seq(min(H29$Year), max(H29$Year), length.out=50))

terms_new<-predict.gam(modH29_GAM_TS$gam, type="terms", newdata=new, se.fit=TRUE)
#terms_new1<-predict.gam(N34_mod$gam, type="response", exclude=c("te(year)", "te(decimal_during)"), newdata=new, se.fit=TRUE)

new$terms<-terms_new$fit[,1]+attr(terms29, "constant")
new$coeff<-terms_new$fit[,2]

new$cont_month=new$dec.during*12+1
new$Month=floor(new$cont_month)
new$coeff.upper<-new$coeff+1.96*terms_new$se.fit[,2]
new$coeff.lower<-new$coeff-1.96*terms_new$se.fit[,2]

new%>%
ggplot(aes(x=Year, y=cont_month, z=coeff)) +
  stat_contour_filled()+
  xlab("Year")+
  theme_bw()+
  scale_y_continuous(name="Month", breaks=c(1,3,5,7,9,11))+
  labs(fill = "c-q slope")+
  theme(text= element_text(size = 14), legend.position="bottom", axis.title.x = element_text(size = 16), axis.title.y = element_text(size = 16))
```
##################
Graphical abstract
##################


```{r}

 new%>%
     ggplot(aes(x=Year, y=cont_month, z=coeff)) +
     stat_contour_filled()+
     xlab("Year")+
   theme_bw()+
     scale_y_continuous(name="Month", breaks=c(1,3,5,7,9,11))+
     labs(fill = "")+
     theme(text= element_text(size = 26), legend.position="bottom", axis.title.x = element_text(size = 30), axis.title.y = element_text(size = 30))+
 labs(caption="Concentration-discharge slope for total nitrogen, catchment H29") +
  theme(plot.caption=element_text(size=30, hjust=0.5, margin=ggplot2::margin(15,0,0,0)))+
   guides(fill=guide_legend(nrow=1))



```

###########################################################
Figure 5, right: Plot of two selected month for illustration
###########################################################

Extract data in February and August using the continuous within-year variable cont_month. Plot the time-varying c-q slopes. 

```{r}
new$cont_month=new$dec.during*12+1

new%>%filter(cont_month>8.4 & cont_month<8.6)%>%
  ggplot( aes(Year, coeff,coeff.lower,coeff.upper)) +
  geom_line(aes(y = coeff), color="red", lwd=2, linetype = "dashed")+
  geom_line(aes(y = coeff.lower))+
  geom_line(aes(y = coeff.upper))+
  ylab("c-q slope")+
  geom_ribbon(aes(ymin=coeff.lower, ymax=coeff.upper),alpha=0.09, fill = "blue")+
  theme_bw()+
geom_line(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(y = coeff), color="blue", lwd=2)+
    geom_line(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(y = coeff.lower))+
  geom_line(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(y = coeff.upper))+
  geom_ribbon(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(ymin=coeff.lower, ymax=coeff.upper),alpha=0.09, fill = "blue")+
  theme(text= element_text(size = 16))
```

