---
title: "Analysis of H29, section 3.3"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load necessary packages
```{r}
library(tidyverse)
library(mgcv)
library(lubridate)
library(readxl)
```

A function to change decimal comma to decimal point
```{r}
replace_comma <- function(values){
    values_parsed <- values %>% as.character() %>% parse_number(locale = locale(decimal_mark = ","))
}
```

Read concentration data för H29, replace decimal comma with decimal point, log-transform total phosphorus, create variables year, month and day
```{r}
mtransp<-read_excel("Analysvärden_220615.xls", sheet="Analysvärden", skip=2, guess_max=20000)%>%filter(ID=="H29", Provtyp=="Manuellt prov")%>%
  mutate_at(c( "Totalkväve_mg/l", "Totalfosfor_mg/l", "Nitrat+nitritkväve_mg/l"),replace_comma)%>%
  mutate(logNO3=log10(`Totalkväve_mg/l`), logTP=log10(`Totalfosfor_mg/l`), logNO3=log10(`Nitrat+nitritkväve_mg/l`), Year=as.numeric(substr(Datum, 1,4)), Month=as.numeric(substr(Datum, 6,7)), Day=as.numeric(substr(Datum, 9,10)))%>%filter(Year<2021)
```

Read discharge data, replace decimal comma with decimal point, create varialbe year, month and day, remove observations with a discharge of 0. 
```{r}
disch<-read_excel("Vattenföring_H-län_210323.xls", sheet="Vattenföring", skip=2, guess_max=100000)%>%filter(ID=="H29")%>%
   mutate_at(c("Dygnsmedelflöde_l/s"),replace_comma)%>%
  mutate(logq=log10(`Dygnsmedelflöde_l/s`), Year=as.numeric(substr(Datum, 1, 4)), Month=as.numeric(substr(Datum, 6,7)), Day=as.numeric(substr(Datum, 9,10)))%>%
  filter(!(`Dygnsmedelflöde_l/s`==0), Year<2021)
```

Combine concentration and discharge data, create a decimal date variable and a variable indicating date within the year.
```{r}
H29<-mtransp%>%left_join(disch)%>%mutate(Date=as.Date( paste(Month, Day , Year , sep = "." )  , format = "%m.%d.%Y"), dec.date=decimal_date(Date), dec.during=dec.date-year(Date))%>%drop_na(logNO3, logq)

```
Table 3:
modH29_constant_withouttrend: model with constant c-q slope and without concentration trend (LM_Constant)
modH29_constant_withtrend: model with constant c-q slope including a concentration trend (GAM_constant)

```{r}
modH29_constant_withouttrend<-lm(logNO3~logq, data=H29)
summary(modH29_constant_withouttrend)

modH29_constant_withtrend<-gamm(logNO3~s(dec.date, bs="ts")+logq, method="REML", data=H29)
summary(modH29_constant_withtrend$gam)
```


Time-varying c-q slope model, changing over seasons (GAM_timevarying and seasonal)

```{r}
H29$logq_cent=H29$logq-mean(H29$logq, na.rm=TRUE)
modH29_GAM_TS<-gamm(logNO3~te(Year, dec.during, bs=c("ts", "cc")) +te(Year, dec.during, by=logq_cent,  bs=c("ts", "cc")),correlation=corCAR1(form=~dec.date),data=H29, method="REML")

summary(modH29_GAM_TS$gam)

```


#########################################################
Figure 5, left: Plot of the year-season varying c-q slope
#########################################################

Define a grid for plotting, years from start to end, season as decimal 0 to 1, logq equal to 1 to extract the slope parameter. Estimated coefficients on this grid and plot
```{r}
new<-expand.grid(dec.during=seq(0,1, length.out=50), logq_cent=rep(1, 50), Year=seq(min(H29$Year), max(H29$Year), length.out=50))

terms_new<-predict.gam(modH29_GAM_TS$gam, type="terms", newdata=new, se.fit=TRUE)
#terms_new1<-predict.gam(N34_mod$gam, type="response", exclude=c("te(year)", "te(decimal_during)"), newdata=new, se.fit=TRUE)

new$terms<-terms_new$fit[,1]+attr(terms_new, "constant")
new$coeff<-terms_new$fit[,2]

new$cont_month=new$dec.during*12+1
new$Month=floor(new$cont_month)
new$coeff.upper<-new$coeff+1.96*terms_new$se.fit[,2]
new$coeff.lower<-new$coeff-1.96*terms_new$se.fit[,2]



breaks <- c(0.15,0.3 ,0.45, 0.6, 0.75, 0.9, 1.05, 1.3, 1.45, 1.6)
Figure5left<-new%>%
ggplot(aes(x=Year, y=cont_month, z=coeff)) +
  stat_contour_filled(breaks=breaks)+
  xlab("Year")+
  theme_bw()+
  scale_y_continuous(name="Month", breaks=c(1,3,5,7,9,11))+
  labs(fill = "c-q slope")+
  theme(text= element_text(size = 12), legend.position="bottom", axis.title.x = element_text(size = 16), axis.title.y = element_text(size = 16))+
  guides(fill=guide_legend(nrow=3))
```

```{r}
ggsave(filename = "Figure 5 left.jpg",    
       plot = Figure5left,
       #plot = last_plot(),
       width = 6, height = 6, device = "jpg", dpi = 300 )
```


##################
Graphical abstract
##################


```{r}
breaks <- c(0, 0.15,0.3 ,0.45, 0.6, 0.75, 0.9, 1.05, 1.3, 1.45, 1.6)
 new%>%
     ggplot(aes(x=Year, y=cont_month, z=coeff)) +
     stat_contour_filled(breaks=breaks)+
     xlab("Year")+
   theme_bw()+
     scale_y_continuous(name="Month", breaks=c(1,3,5,7,9,11))+
     labs(fill = "")+
     theme(text= element_text(size = 26), legend.position="bottom", axis.title.x = element_text(size = 30), axis.title.y = element_text(size = 30))+
 labs(caption="Concentration-discharge slope for nitrate, catchment H29") +
  theme(plot.caption=element_text(size=30, hjust=0.5, margin=ggplot2::margin(15,0,0,0)))+
   guides(fill=guide_legend(nrow=2))



```

###########################################################
Figure 6, right: Pot of two selected month for illustration
###########################################################

Sort out data in August using the continuous within year variable. Extract data in February and August

```{r}
new$cont_month=new$dec.during*12+1

Figure5right<-new%>%filter(cont_month>8.4 & cont_month<8.6)%>%
  ggplot( aes(Year, coeff,coeff.lower,coeff.upper)) +
  geom_line(aes(y = coeff), color="red", lwd=2, linetype = "dashed")+
  geom_line(aes(y = coeff.lower))+
  geom_line(aes(y = coeff.upper))+
  ylab("c-q slope")+
  geom_ribbon(aes(ymin=coeff.lower, ymax=coeff.upper),alpha=0.09, fill = "blue")+
  theme_bw()+
geom_line(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(y = coeff), color="blue", lwd=2)+
    geom_line(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(y = coeff.lower))+
  geom_line(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(y = coeff.upper))+
  geom_ribbon(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(ymin=coeff.lower, ymax=coeff.upper),alpha=0.09, fill = "blue")+
  theme(text= element_text(size = 16))
```

```{r}
ggsave(filename = "Figure 5 right.jpg",    
       plot = Figure5right,
       #plot = last_plot(),
       width = 6, height = 6, device = "jpg", dpi = 300 )
```

model with no temporal trend term estimated.
```{r}
# H29_a<-H29%>%drop_na(logNO3, logq)
# H29_a$logq_cent=H29_a$logq-mean(H29_a$logq)
# modH29_GAM_TS_utan_trend<-gamm(logNO3~te(Year, dec.during, by=logq_cent,  bs=c("ts", "cc")),correlation=corCAR1(form=~dec.date),data=H29_a, method="REML")
# new<-expand.grid(dec.during=seq(0,1, length.out=50), logq_cent=rep(1, 50), Year=seq(min(H29$Year), max(H29$Year), length.out=50))
# 
# terms_new<-predict.gam(modH29_GAM_TS_utan_trend$gam, type="terms", newdata=new, se.fit=TRUE)
# #terms_new1<-predict.gam(N34_mod$gam, type="response", exclude=c("te(year)", "te(decimal_during)"), newdata=new, se.fit=TRUE)
# 
# #new$terms<-terms_new$fit[,1]+attr(terms_new, "constant")
# new$coeff<-terms_new$fit[,1]
# 
# new$cont_month=new$dec.during*12+1
# new$Month=floor(new$cont_month)
# new$coeff.upper<-new$coeff+1.96*terms_new$se.fit[,1]
# new$coeff.lower<-new$coeff-1.96*terms_new$se.fit[,1]
# 
# 
# 
# breaks <- c(0.05,0.2 ,0.35, 0.5, 0.65, 0.8, 0.95, 1.1, 1.25, 1.4, 1.55, 1.7, 1.95)
# new%>%
# ggplot(aes(x=Year, y=cont_month, z=coeff)) +
#   stat_contour_filled(breaks=breaks)+
#   xlab("Year")+
#   theme_bw()+
#   scale_y_continuous(name="Month", breaks=c(1,3,5,7,9,11))+
#   labs(fill = "c-q slope")+
#   theme(text= element_text(size = 14), legend.position="bottom", axis.title.x = element_text(size = 16), axis.title.y = element_text(size = 16))+
#   guides(fill=guide_legend(nrow=3))
# 
# new$cont_month=new$dec.during*12+1
# 
# new%>%filter(cont_month>8.4 & cont_month<8.6)%>%
#   ggplot( aes(Year, coeff,coeff.lower,coeff.upper)) +
#   geom_line(aes(y = coeff), color="red", lwd=2, linetype = "dashed")+
#   geom_line(aes(y = coeff.lower))+
#   geom_line(aes(y = coeff.upper))+
#   ylab("c-q slope")+
#   geom_ribbon(aes(ymin=coeff.lower, ymax=coeff.upper),alpha=0.09, fill = "blue")+
#   theme_bw()+
# geom_line(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(y = coeff), color="blue", lwd=2)+
#     geom_line(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(y = coeff.lower))+
#   geom_line(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(y = coeff.upper))+
#   geom_ribbon(data=new%>%filter(cont_month>2.4 & cont_month<2.6),aes(ymin=coeff.lower, ymax=coeff.upper),alpha=0.09, fill = "blue")+
#   theme(text= element_text(size = 16))
```

################################################
Figure S3: Plot trend estimate in modH29_GAM_TS:
################################################
```{r}
new<-expand.grid(dec.during=seq(0,1, length.out=50), logq_cent=rep(1, 50), Year=seq(min(H29$Year), max(H29$Year), length.out=50))

terms_new<-predict.gam(modH29_GAM_TS$gam, type="terms", newdata=new, se.fit=TRUE)
#terms_new1<-predict.gam(N34_mod$gam, type="response", exclude=c("te(year)", "te(decimal_during)"), newdata=new, se.fit=TRUE)

#new$terms<-terms_new$fit[,1]+attr(terms_new, "constant")
new$trend<-terms_new$fit[,1]+attr(terms_new, "constant")

new$cont_month=new$dec.during*12+1
new$Month=floor(new$cont_month)
new$trend.upper<-new$trend+1.96*terms_new$se.fit[,1]
new$trend.lower<-new$trend-1.96*terms_new$se.fit[,1]
new$coeff<-terms_new$fit[,2]

new$coeff.upper<-new$coeff+1.96*terms_new$se.fit[,2]
new$coeff.lower<-new$coeff-1.96*terms_new$se.fit[,2]



breaks <- c(0.05,0.2 ,0.35, 0.5, 0.65, 0.8, 0.95, 1.1, 1.25, 1.4, 1.55, 1.7, 1.95)
FigureS3<-new%>%
ggplot(aes(x=Year, y=cont_month, z=trend)) +
  stat_contour_filled()+
  xlab("Year")+
  theme_bw()+
  scale_y_continuous(name="Month", breaks=c(1,3,5,7,9,11))+
  labs(fill = "Estimated nitrate levels")+
  theme(text= element_text(size = 14), legend.position="bottom", axis.title.x = element_text(size = 16), axis.title.y = element_text(size = 16))+
  guides(fill=guide_legend(nrow=2))+
  labs(fill="")+
 #  labs(caption="Estimated NO3 levels")+
 theme(plot.caption=element_text(size=16, hjust=0.5, margin=ggplot2::margin(15,0,0,0)))
```

```{r}
ggsave(filename = "Figure S3.jpg",    
       plot = FigureS3,
       #plot = last_plot(),
       width = 6, height = 6, device = "jpg", dpi = 300 )
```

#################################################################
Figure S4: Prediction plot
##################################################################
```{r}
modH29_GAM_TS_constant_cq<-gamm(logNO3~te(Year, dec.during, bs=c("ts", "cc")) +logq_cent,correlation=corCAR1(form=~dec.date),data=H29, method="REML")
modH29_GAM_TS<-gamm(logNO3~te(Year, dec.during, bs=c("ts", "cc")) +te(Year, dec.during, by=logq_cent,  bs=c("ts", "cc")),correlation=corCAR1(form=~dec.date),data=H29, method="REML")


H29$resid_const<-residuals(modH29_GAM_TS_constant_cq$gam)
H29$resid_TS<-residuals(modH29_GAM_TS$gam)
H29$pred_const<-as.numeric(predict(modH29_GAM_TS_constant_cq$gam , type="response"))
H29$pred_TS<-as.numeric(predict(modH29_GAM_TS$gam, type="response"))


H29%>%pivot_longer(cols=c(pred_TS, pred_const))->H29_b

labels<- as_labeller(c(pred_const="GAM[constant]",pred_TS="GAM[timevarying~and~seasonal]"),default=label_parsed)


FigureS4bottom<-H29_b%>%
  ggplot(aes(x=logNO3,y=value))+
  geom_point(col="blue")+
  geom_abline(slope=1, intercept=0, lwd=2)+
  facet_wrap(~name, labeller=labels)+
  ylab("Predicted values")+
  xlab("Nitrate (log-transformed)")+
  theme_classic()+
   theme(strip.text.x = element_text(size = 16),text= element_text(size = 16))



```


```{r}
ggsave(filename = "Figure S4 bottom.jpg",    
       plot = FigureS4bottom,
       #plot = last_plot(),
       width = 7, height = 5, device = "jpg", dpi = 300 )

```


