---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load necessary packages

```{r}
library(tidyverse)
library(mgcv)
library(lubridate)
library(readxl)
```

To read data from files decimal commas need to be replaces with decimal points. The following function can be used for this.
```{r}
replace_comma <- function(values){
    values_parsed <- values %>% as.character() %>% parse_number(locale = locale(decimal_mark = ","))
}

```


Read concentration values for station O14, replace decimal commas, log-transform variables and create new variables that represent year, month and day.
```{r}
mtransp<-read_excel("Analysvärden_220615.xls", sheet="Analysvärden", skip=2, guess_max=20000)%>%filter(ID=="O14", Provtyp=="Manuellt prov")%>%
  mutate_at(c( "Totalkväve_mg/l", "Totalfosfor_mg/l", "Nitrat+nitritkväve_mg/l"),replace_comma)%>%
  mutate(logTN=log10(`Totalkväve_mg/l`), logTP=log10(`Totalfosfor_mg/l`),logNO3=log10(`Nitrat+nitritkväve_mg/l`), Year=as.numeric(substr(Datum, 1,4)), Month=as.numeric(substr(Datum, 6,7)), Day=as.numeric(substr(Datum, 9,10)))%>%
  filter(Year<2021)%>%
  mutate(logNO3=case_when(`Nitrat+nitritkväve_mg/l`==0~log10(0.001),
                          TRUE~logNO3))
```

Read discharge values for station O14, replace decimal commas, log-transform variables and create new variables that represent year, month and day.Remove days with a discharge equal to 0.
```{r}
disch<-read_excel("Vattenföring_O14_220615.xls", sheet="Vattenföring", skip=2, guess_max=100000)%>%filter(ID=="O14")%>%
   mutate_at(c("Dygnsmedelflöde_l/s"),replace_comma)%>%
  mutate(logq=log10(`Dygnsmedelflöde_l/s`), Year=as.numeric(substr(Datum, 1, 4)), Month=as.numeric(substr(Datum, 6,7)), Day=as.numeric(substr(Datum, 9,10)))%>%
  filter(!(`Dygnsmedelflöde_l/s`==0), Year<2021)
```

Combine concentration and discharge data sets, create a date variable, a variable giving date in decimal form and a variable giving time within the year. An additional vary small discharge value is set to missing.
```{r}
O14<-mtransp%>%left_join(disch)%>%mutate(Date=as.Date( paste(Month, Day , Year , sep = "." )  , format = "%m.%d.%Y"), dec.date=decimal_date(Date), dec.during=dec.date-year(Date))%>%drop_na(logNO3)


#O14$logq[O14$Datum==as.Date("2018-10-01")]<-NA

```

##############################################################################
Section 3.2: Estimating a constant c-q relationship in the presence of a trend
##############################################################################

modO14 is a model that estimates a constant c-q slope while a trend in log TN is also included (GAM_Constant)
modO14_notrend is the same model without trend in concentration (LM_constant)

```{r}
#modO14<-gamm(logNO3~s(dec.date, bs="ts")+logq,correlation=corCAR1(form = ~dec.date),method="REML", data=O14)
#summary(modO14$gam)

modO14<-gamm(logNO3~s(dec.date, bs="ts")+logq,method="REML", data=O14)
summary(modO14$gam)

#modO14_notrend<-gls(logNO3~logq, correlation=corCAR1(form = ~dec.date),data=O14, na.action=na.omit)
#summary(modO14_notrend)

modO14_notrend<-lm(logNO3~logq, data=O14, na.action=na.omit)
summary(modO14_notrend)


O14$logq_cent<-O14$logq-mean(O14$logq, na.rm=TRUE)

modO14_notrend<-lm(logNO3~logq_cent, data=O14, na.action=na.omit)
summary(modO14_notrend)

modO14<-gamm(logNO3~s(dec.date, bs="ts")+logq_cent,method="REML", data=O14)
summary(modO14$gam)
```
####################################################################################
Figure S.3: Plot of the trend in concentration, no relation to discharge is modelled
####################################################################################

```{r}
modO14_trend_only<-gamm(logNO3~s(dec.date, bs="ts"),correlation=corCAR1(form = ~dec.date),method="REML", data=O14)
terms<-predict.gam(modO14_trend_only$gam, type="terms", newdata=O14, se.fit=TRUE)

O14$fit<-terms$fit[,1]+attr(terms, "constant")

ggplot(O14, aes(dec.date, logNO3,fit)) +
  geom_point(aes(x=dec.date, y=logNO3))+
  geom_line(aes(y = fit), color="red", lwd=2)+
  ylab("Nitrate (log-transformed)")+
  xlab("Date")+
  theme_bw()+
  theme(text= element_text(size = 16))
```


###############################################
Section 3.3: Estimating nonlinear relationships
###############################################

modO14_GAM_N: Following a GAM_Nonlinear model, a nonlinear c-q relationship as well as a trend in concentration levels.

```{r}
modO14_GAM_N<-gamm(logNO3~s(dec.date, bs="ts")+s(logq),correlation=corCAR1(form = ~dec.date),method="REML", data=O14)
summary(modO14_GAM_N$gam)

```

Extract the fit and confidence band for the nonlinear relationship
```{r}
terms<-predict(modO14_GAM_N$gam, type="terms", newdata=O14, se.fit=TRUE)
O14$sereg<-terms$se.fit[,2]
O14$line<-terms$fit[,2]+mean(terms$fit[,1], na.rm=TRUE)+attr(terms, "constant")
O14$Line_upper<-O14$line+2*terms$se.fit[,2]
O14$Line_lower<-O14$line-2*terms$se.fit[,2]
```

##################################################
Figure 2, left: Nonlinear c-q relationships in O14
##################################################
```{r}
O14%>%
  ggplot(aes(x=logq, y=logNO3, Line_lower,Line_upper))+
  geom_point()+
  xlab("Discharge (log-transformed)")+
  ylab("Nitrate (log-transformed)")+
  geom_ribbon(aes(ymin=Line_lower, ymax=Line_upper), fill="lightgreen", alpha=0.5)+
    geom_line(aes(x=logq, y=line), col="darkgreen",lwd=1)+
  theme_bw()+
  theme(text= element_text(size = 16))
```

Extract the nonlinear c-q slope estimate using derivatives
```{r}
library(gratia)
ff<-derivatives(modO14_GAM_N)
```

###########################################
Figure 2, right: Nonlinear c-q slope in O14
###########################################
```{r}
ff%>%filter(var=="logq")%>%
  ggplot(aes(data, derivative, lower, upper))+
  geom_line(aes(y=derivative), col="red", lwd=2)+
  geom_line(aes(y=lower))+
  geom_line(aes(y=upper))+
  ylab("c-q slope")+
  xlab("Discharge (log-transformed)")+
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.09, fill="blue")+
  theme_bw()+
   theme(text= element_text(size = 16))

   
```

Seasonailty in residuals?
```{r}
O14$cont_month=O14$dec.during*12+1
O14%>%drop_na(logNO3, logq)->O14_a

O14_a$resid<-residuals(modO14_GAM_N$gam)

resid.points<-ggplot(O14_a, aes(x=Year,y=cont_month))+geom_point(aes(colour=resid), lwd=4)+scale_colour_gradient2(low = "darkblue", mid = "white", high = "darkred", midpoint = 0, space = "rgb", na.value = "grey50", guide = "colourbar", guide_legend(title="Residuals"))+
  theme_classic()+
    ylab("")+
  xlab("")+
  theme(panel.background = element_rect(fill = "gray56",
                                colour = "gray56",
                                size = 0.5, linetype = "solid"))
resid.points

dists <- as.matrix(dist(cbind(O14_a$Year, O14_a$cont_month)))

dists.inv <- 1/dists
diag(dists.inv) <- 0
 
dists.inv[1:5, 1:5]
 
library(ape)

Moran.I(O14_a$resid, dists.inv)

```





Estimating nonlinear c-q relationships for different seasons

Define seasons
```{r}
O14$seas<-1
O14$seas[O14$Month==3| O14$Month==4 | O14$Month ==5]<-2
O14$seas[O14$Month==6| O14$Month==7 | O14$Month ==8]<-3
O14$seas[O14$Month==9| O14$Month==10 | O14$Month ==11]<-4
O14$seas<-as.factor(O14$seas)
```

modO14_GAM_NS following a GAM_Nonlinear and seasonal model with nonlinear relationships for separate seasons
```{r}
modO14_GAM_NS<-gamm(logNO3~s(dec.date, bs="ts")+seas+s(logq, by=seas),correlation=corCAR1(form = ~dec.date),method="REML", data=O14)
summary(modO14_GAM_NS$gam)
```

Extract fits and confidence bands for the seasonal c-q slopes models, March to May
```{r}
terms1<-predict(modO14_GAM_NS$gam, type="terms", newdata=O14, se.fit=TRUE)


O14$sereg<-terms1$se.fit[,4]
O14$line<-terms1$fit[,1]+terms1$fit[,4]+mean(terms1$fit[,2], na.rm=TRUE)+attr(terms1, "constant")
O14$Line_upper<-O14$line+2*terms1$se.fit[,4]
O14$Line_lower<-O14$line-2*terms1$se.fit[,4]
```

###########################################################
Figure 3, upper row, left: Nonlinear relationship in spring
###########################################################
```{r}
O14%>%
  filter(seas==2)%>%
  ggplot(aes(x=logq, y=logNO3, Line_lower,Line_upper))+
  geom_point()+
  xlab("Discharge (log-transformed)")+
  ylab("Nitrate (log-transformed)")+
  geom_ribbon(aes(ymin=Line_lower, ymax=Line_upper), fill="lightgreen", alpha=0.5)+
    geom_line(aes(x=logq, y=line), col="darkgreen",lwd=1)+
  theme_bw()+
  xlim(-1,3.2)+
 theme(text= element_text(size = 16), plot.title = element_text(size=16))+
  labs(title="Spring")
```

Extract fits and confidence bands for the seasonal c-q slopes models, June to August
```{r}
O14$sereg<-terms1$se.fit[,5]
O14$line<-terms1$fit[,1]+terms1$fit[,5]+mean(terms1$fit[,2], na.rm=TRUE)+attr(terms1, "constant")
O14$Line_upper<-O14$line+2*terms1$se.fit[,5]
O14$Line_lower<-O14$line-2*terms1$se.fit[,5]
```

#############################################################
Figure 3, upper row, middle: Nonlinear relationship in Summer
#############################################################
```{r}
O14%>%
  filter(seas==3)%>%
  ggplot(aes(x=logq, y=logNO3, Line_lower,Line_upper))+
  geom_point()+
  xlab("Discharge (log-transformed)")+
  ylab("Nitrate (log-transformed)")+
  geom_ribbon(aes(ymin=Line_lower, ymax=Line_upper), fill="lightgreen", alpha=0.5)+
    geom_line(aes(x=logq, y=line), col="darkgreen",lwd=1)+
  theme_bw()+
   theme(text= element_text(size = 16), plot.title = element_text(size=16))+
  labs(title="Summer")
```

Extract fits and confidence bands for the seasonal c-q slopes models, September to November
```{r}

O14$sereg<-terms1$se.fit[,6]
O14$line<-terms1$fit[,1]+terms1$fit[,6]+mean(terms1$fit[,2], na.rm=TRUE)+attr(terms1, "constant")
O14$Line_upper<-O14$line+2*terms1$se.fit[,6]
O14$Line_lower<-O14$line-2*terms1$se.fit[,6]
```

############################################################
Figure 3, upper row, right: Nonlinear relationship in autumn
############################################################
```{r}
O14%>%
  filter(seas==4)%>%
  ggplot(aes(x=logq, y=logNO3, Line_lower,Line_upper))+
  geom_point()+
  xlab("Discharge (log-transformed)")+
  ylab("Nitrate (log-transformed)")+
  geom_ribbon(aes(ymin=Line_lower, ymax=Line_upper), fill="lightgreen", alpha=0.5)+
    geom_line(aes(x=logq, y=line), col="darkgreen",lwd=1)+
  theme_bw()+
   theme(text= element_text(size = 16), plot.title = element_text(size=16))+
  labs(title="Autumn")
```

Extract fits and confidence bands for the seasonal c-q slopes models, December to February
```{r}

O14$sereg<-terms1$se.fit[,3]
O14$line<-terms1$fit[,1]+terms1$fit[,3]+mean(terms1$fit[,2], na.rm=TRUE)+attr(terms1, "constant")
O14$Line_upper<-O14$line+2*terms1$se.fit[,3]
O14$Line_lower<-O14$line-2*terms1$se.fit[,3]
```

Not shown in article, Figure December to February
```{r}
O14%>%
  filter(seas==1)%>%
  ggplot(aes(x=logq, y=logNO3, Line_lower,Line_upper))+
  geom_point()+
  xlab("Discharge (log-transformed)")+
  ylab("Total nitrogen (log-transformed)")+
  geom_ribbon(aes(ymin=Line_lower, ymax=Line_upper), fill="lightgreen", alpha=0.5)+
    geom_line(aes(x=logq, y=line), col="darkgreen",lwd=1)+
  theme_bw()+
   theme(text= element_text(size = 16), plot.title = element_text(size=16))+
  labs(title="Winter")

```


Extract c-q slopes from the seasonal model using derivatives 
```{r}
termsseas<-predict.gam(modO14_GAM_NS$gam, type="terms", newdata=O14, se.fit=TRUE)

ff_seas<-derivatives(modO14_GAM_NS)
```

########################################################
Figure 3, lower row, left: Nonlinear c-q slope in spring
########################################################
```{r}
ff_seas%>%filter(var=="logq", smooth=="s(logq):seas2")%>%
  ggplot(aes(data, derivative, lower, upper))+
  geom_line(aes(y=derivative), col="red", lwd=2)+
  geom_line(aes(y=lower))+
  geom_line(aes(y=upper))+
  ylab("c-q slope")+
  xlab("Discharge (log-transformed)")+
  ylim(-0.85, 4.5)+
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.09, fill="blue")+
  theme_bw()+
  theme(text= element_text(size = 16), plot.title = element_text(size=16))+
  labs(title="Spring")
```

##########################################################
Figure 3, lower row, middle: Nonlinear c-q slope in summer
##########################################################
```{r}
ff_seas%>%filter(var=="logq", smooth=="s(logq):seas3")%>%
  ggplot(aes(data, derivative, lower, upper))+
  geom_line(aes(y=derivative), col="red", lwd=2)+
  geom_line(aes(y=lower))+
  geom_line(aes(y=upper))+
  ylab("c-q slope")+
  xlab("Discharge (log-transformed)")+
    ylim(-0.85, 4.5)+
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.09, fill="blue")+
  theme_bw()+
  theme(text= element_text(size = 16), plot.title = element_text(size=16))+
  labs(title="Summer")
```

#########################################################
Figure 3, lower row, right: Nonlinear c-q slope in autumn
#########################################################
```{r}
ff_seas%>%filter(var=="logq", smooth=="s(logq):seas4")%>%
  ggplot(aes(data, derivative, lower, upper))+
  geom_line(aes(y=derivative), col="red", lwd=2)+
  geom_line(aes(y=lower))+
  geom_line(aes(y=upper))+
  ylab("c-q slope")+
  xlab("Discharge (log-transformed)")+
  ylim(-0.85, 4.5)+
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.09, fill="blue")+
  theme_bw()+
  theme(text= element_text(size = 16), plot.title = element_text(size=16))+
  labs(title="Autumn")
```

Figure December to February (not shown in article)
```{r}
ff_seas%>%filter(var=="logq", smooth=="s(logq):seas1")%>%
  ggplot(aes(data, derivative, lower, upper))+
  geom_line(aes(y=derivative), col="red", lwd=2)+
  geom_line(aes(y=lower))+
  geom_line(aes(y=upper))+
  ylab("c-q slope")+
  xlab("Discharge (log-transformed)")+
  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.09, fill="blue")+
  theme_bw()+
    theme(text= element_text(size = 16), plot.title = element_text(size=16))+
  labs(title="Winter")


```

```{r}
gam.check(modO14_GAM_NS$gam)
gam.check(modO14_GAM_N$gam)
```

Figure S4: Comparison of observed and predicted values using a seasonal constant slope model and the GAM_nonlinear_and_seasonal model.
```{r}
modO14_GAM_NS_constant_cq<-gamm(logNO3~s(dec.date, bs="ts")+seas+logq,correlation=corCAR1(form = ~dec.date),method="REML", data=O14)

O14_a<-O14%>%drop_na(logNO3, logq)

O14_a$resid_const<-residuals(modO14_GAM_NS_constant_cq$gam)
O14_a$resid_TS<-residuals(modO14_GAM_NS$gam)
O14_a$pred_const<-as.numeric(predict(modO14_GAM_NS_constant_cq$gam , type="response"))
O14_a$pred_NS<-as.numeric(predict(modO14_GAM_NS$gam, type="response"))


O14_a%>%pivot_longer(cols=c(pred_NS, pred_const))->O14_b

labels<- as_labeller(c(pred_const="GAM[constant]",pred_NS="GAM[nonlinear~and~seasonal]"),default=label_parsed)


O14_b%>%
  ggplot(aes(x=logNO3,y=value))+
  geom_point(col="blue")+
  geom_abline(slope=1, intercept=0, lwd=2)+
  facet_wrap(~name, labeller=labels)+
  ylab("Predicted values")+
  xlab("Nitrate (log-transformed)")+
  theme_classic()+
   theme(strip.text.x = element_text(size = 16), text= element_text(size = 16))


```


