---
title: "Simulations, Section 3.2"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Install and load necessary packages.
```{r}
library(tidyverse)
library(broom)
library(mgcv)
```

########################
Simulations, section 3.1
########################


Figure: An example series (not shown in paper)

```{r}
set.seed(34)
r<-rnorm(100, 0, sqrt(1))
t<-seq(1:100)
q<-rnorm(100, 0, sqrt(1))
c=0+0.3*q+r
c1<-c
c1[1:100]<-c[1:100]+seq(1:100)/50
datatest<-data.frame(q=q, c=c, t=t, c1=c1)

datatest%>%
  ggplot(aes(x=t, y=c1))+
  geom_point()+
  geom_line(aes(x=t, y=q), col="blue")+
  ylab("")+
  xlab("Time")+
  theme_bw()+
  theme(text= element_text(size = 20))
```

Simdata - Function to create generated data with the desired distributions and values for b
variable c does not contain a trend, while c1 is the same series including a trend

Simdata_q - Simulate data with a trend in discharge. 
```{r}
N=1000
div=100


set.seed(34)
Simdata<-function(n=n, d=100, g=1, beta=0.3, div=10){
	r<-rnorm(d, 0, sqrt(1))
t<-seq(1:d)
q<-rnorm(d, 0, sqrt(1))
c<-0+beta*q+r
c1<-c
c1[g:d]<-c[g:d]+seq(1:((d-g)+1))/div
sim<-rep(n, d)
out<-data.frame(q,c, c1, t,sim)

}


Simdata_q<-function(n=n, d=100, g=1, beta=0.3, div=10){
	r<-rnorm(d, 0, sqrt(1))
t<-seq(1:d)
q<-rnorm(d, 0, sqrt(1))
c<-0+beta*q+r
q1<-q
q1[g:d]<-q[g:d]+seq(1:((d-g)+1))/div
c2<-0+beta*q1+r
sim<-rep(n, d)
out<-data.frame(q,c, c2, q1, t,sim)

}
```

Simulate data
```{r}
s1 <- data.frame(q=numeric(),c=numeric(), c1=numeric(), t=numeric(), sim=numeric()) 
for (n in 1:N){
  s<-Simdata(n)
  s1<-full_join(s1,s)
}
```

Run constant c-q slope models (LM_constant) for data without (c) and with (c1) trends in concentration
```{r}
s1%>%
  group_by(as.factor(sim))%>%
    nest() %>% 
  mutate(model = map(data, ~lm(c ~ q, data = .x) %>% 
                       tidy)) %>% 
  unnest(model) %>%
  filter(term == 'q')%>% rename(sim=`as.factor(sim)`, beta_lm_no_trend=estimate)%>%select(sim, beta_lm_no_trend) ->a1

s1%>%
group_by(as.factor(sim))%>%
    nest() %>% 
  mutate(model = map(data, ~lm(c1 ~ q, data = .x) %>% 
                       tidy)) %>% 
   unnest(model) %>%
  filter(term == 'q')%>% rename(sim=`as.factor(sim)`, beta_lm_c_trend=estimate)%>%select(sim, beta_lm_c_trend) ->a2
```


Functions to run constant slope models including trend (GAM_constant) on data without trend (c) and with trend (c1), and using a discharge variable without trend (q) or one with trend (q1).
```{r}
run_mod_c1<-function(data=data){
mm <- gamm(c1 ~  s(t, bs="ts")+q , data=data,method="REML")  
	ss<-summary(mm$gam)
	return(ss$p.coeff[2])
}	

run_mod_c<-function(data=data){
mm <- gamm(c ~  s(t, bs="ts")+q , data=data,method="REML")  
	ss<-summary(mm$gam)
	return(ss$p.coeff[2])
}	

run_mod_q1<-function(data=data){
mm <- gamm(c2 ~  s(t, bs="ts")+q1 , data=data,method="REML")  
	ss<-summary(mm$gam)
	return(ss$p.coeff[2])
}	

extract_trend<-function(q=q, c=c, t=t, data=data){
mm <- gamm(c ~  s(t, bs="ts")+q , data=data,method="REML")  
	return(predict(mm$gam, type="terms")[,2])
}
```

Additional functions to extract edf from models
```{r}
run_mod_edf_c<-function(data=data){
mm <- gamm(c ~  s(t, bs="ts")+q , data=data,method="REML")  
	ss<-summary(mm$gam)
	return(ss$edf)
}	

run_mod_edf_c1<-function(data=data){
mm <- gamm(c1 ~  s(t, bs="ts")+q , data=data,method="REML")  
	ss<-summary(mm$gam)
	return(ss$edf)
}	

run_mod_edf_q1<-function(data=data){
mm <- gamm(c2 ~  s(t, bs="ts")+q1 , data=data,method="REML")  
	ss<-summary(mm$gam)
	return(ss$edf)
}
```

Run constant slope model with concentration trend (GAM_constant) on data without (c) and with (c1) trend in concentration.

```{r}
s1%>%
  group_by(as.factor(sim))%>%
  nest()%>%
  mutate(model =map(data, ~run_mod_c(data=.x)))%>%
  unnest(model)%>% rename(sim=`as.factor(sim)`, beta_gam_no_trend=model)%>%select(sim, beta_gam_no_trend)   ->a3

s1%>%
  group_by(as.factor(sim))%>%
  nest()%>%
  mutate(model =map(data, ~run_mod_c1(data=.x)))%>%
  unnest(model)%>% rename(sim=`as.factor(sim)`, beta_gam_c_trend=model)%>%select(sim, beta_gam_c_trend) ->a4
```

Simulate data with trend in discharge
```{r}
set.seed(34)
s2 <- data.frame(q=numeric(),c=numeric(), c1=numeric(), q1=numeric(), t=numeric(), sim=numeric()) 
for (n in 1:N){
  s<-Simdata_q(n)
  s2<-full_join(s2,s)
}
```

Run constant c-q slope models (LM_constant) for data with trends in discharge (q1) and corresponding concentration(c2) 
```{r}
s2%>%
group_by(as.factor(sim))%>%
    nest() %>% 
  mutate(model = map(data, ~lm(c2 ~ q1, data = .x) %>% 
                       tidy)) %>% 
   unnest(model) %>%
  filter(term == 'q1')%>% rename(sim=`as.factor(sim)`, beta_lm_q_trend=estimate)%>%select(sim, beta_lm_q_trend) ->a5
```

Run constant c-q model with trend (GAM_constant) for data with trend in discharge (q1) and corresponding concentration (c2)
```{r}

s2%>%
  group_by(as.factor(sim))%>%
  nest()%>%
  mutate(model =map(data, ~run_mod_q1(data=.x)))%>%
  unnest(model)%>% rename(sim=`as.factor(sim)`, beta_gam_q_trend=model)%>%select(sim, beta_gam_q_trend)   ->a6


s2%>%
  group_by(as.factor(sim))%>%
  nest()%>%
  mutate(model =map(data, ~extract_trend(c=c2, q=q1, t=t, data=.x)))%>%
  unnest(model)  ->a7
```

#############################################################################################
Figure 1: Boxplots of modelled c-q slopes for the six combination of models and scenarios
#############################################################################################
```{r}
a1%>%left_join(a2, by="sim")%>%left_join(a3, by="sim")%>%left_join(a4,by="sim")%>%left_join(a5,by="sim")%>%left_join(a6,by="sim")->a_all

a_all%>%
  pivot_longer(cols=c("beta_lm_no_trend", "beta_lm_c_trend",  "beta_gam_no_trend", "beta_gam_c_trend", "beta_lm_q_trend", "beta_gam_q_trend"))%>%
  mutate(groups=case_when(name=="beta_lm_no_trend"~"no trend present (Model 1)",
                          name=="beta_lm_c_trend"~"c trend present (Model 1)",
                          name=="beta_gam_no_trend"~"no trend present (Model 2)",
                          name=="beta_gam_c_trend"~"c trend present (Model 2)",
                           name=="beta_lm_q_trend"~"q trend present (Model 1)",
                          name=="beta_gam_q_trend"~"q trend present (Model 2)"))->a_plot
a_plot$groups<-factor(a_plot$groups, levels=c("no trend present (Model 1)", "no trend present (Model 2)", "c trend present (Model 1)", "c trend present (Model 2)", "q trend present (Model 1)",  "q trend present (Model 2)"))

a_plot%>%mutate(trend=case_when(groups=="no trend present (Model 1)"~"No trend simulated (i)",
                                groups=="no trend present (Model 2)"~"No trend simulated (i)",
                                groups=="c trend present (Model 1)"~"Trend in concentration simulated (ii)", 
                                groups=="c trend present (Model 2)"~"Trend in concentration simulated (ii)",
                                groups=="q trend present (Model 1)"~"Trend in discharge simulated (iii)",
                                groups=="q trend present (Model 2)"~"Trend in discharge simulated (iii)"),
                model1=case_when(groups=="no trend present (Model 1)"~"LM",
                                groups=="no trend present (Model 2)"~"GAM",
                                groups=="c trend present (Model 1)"~"LM", 
                                groups=="c trend present (Model 2)"~"GAM", 
                                groups=="q trend present (Model 1)"~"LM",
                                groups=="q trend present (Model 2)"~"GAM"
                                ))%>%
  mutate(model=fct_relevel(model1, levels=c("LM", "GAM")))%>%
  ggplot(aes(y=value, x=model))+
  geom_boxplot()+
  scale_x_discrete(name = "",
                  labels = c(expression("LM"["constant"]),
                             expression("GAM"["constant"])))+
  xlab("")+
  ylab("Slope estimate b")+
  theme_bw()+
     theme(text= element_text(size = 20))+
    facet_wrap(~trend)

```

